Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-15 cpp_paper.tex
+
compiles
revisions up through end of section 3
remove discussion section
remove full soundness section, add bus-draft conclusion

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}
\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage{todonotes}
\usepackage{blindtext}
\usepackage{listings}
\lstset{
  columns=fullflexible,
  numbers=left,
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  morekeywords={mov,add,call},
  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
  emphstyle=\color{green},
  emph={[2]cr3},
  emphstyle={[2]\color{violet}},
  morecomment=[l]{;;}
}
% Hack for Colin to edit in low light: https://tex.stackexchange.com/questions/40495/invert-background-and-text-colours-across-whole-document-with-pdflatex
%Commands definitions
\newcommand{\setbackgroundcolour}{\pagecolor[rgb]{0.19,0.19,0.19}}  
\newcommand{\settextcolour}{\color[rgb]{0.77,0.77,0.77}}    
\newcommand{\invertbackgroundtext}{\setbackgroundcolour\settextcolour}
%Command execution. 
%If this line is commented, then the appearance remains as usual.
%\invertbackgroundtext
\usepackage{wrapfig}
\newtheorem*{remark}{Remark}

% Packages.
\usepackage{amsmath}
% swiped from iris.sty in the iris repo
\newcommand{\wand}{\mathrel{-\!\!\ast}}
\newcommand{\wandIff}{\mathrel{\ast\!\!{-}\!\!\ast}}
% If we need multiple references to a footnote, via \cref:
\usepackage{cleveref}
\crefformat{footnote}{#2\footnotemark[#1]#3}
\usepackage[T1]{fontenc}
\usepackage{fontawesome}
\usepackage[utf8]{inputenc}
\usepackage{iris}
\usepackage{mathpartir}
\renewcommand{\TirNameStyle}[1]{\hypertarget{#1}{\textsc{#1}}}
\newcommand{\RULE}[1]{\hyperlink{#1}{\textsc{#1}}\xspace}
\usepackage{xspace}
\input{macros}

\newcommand{\mytodo}[1]{% <==========================================
  \todo[linecolor=white, bordercolor=white, textcolor=white]{#1}%
}

@@

\todo[inline]{Colin think through instruction pointer updates vs logic,
think about framing plus cr3 update and modality shifts
}

\input{intro.tex}
%\input{motivation.tex}
\input{background.tex}
\input{syntaxsemantics.tex}
\input{programlogic.tex}
\input{soundness.tex}
\input{examples.tex}
\input{discussion.tex}
\input{relwork.tex}
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{vmm}

>>>>

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{amsmath}
%\usepackage{amssymb}

\usepackage{hyperref}
\usepackage{array}
%\usepackage{amsthm}
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{xspace}
%\usepackage{listings}
\usepackage{graphicx}

\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage{todonotes}
\usepackage{blindtext}
\usepackage{listings}
\lstset{
  columns=fullflexible,
  numbers=left,
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  morekeywords={mov,add,call},
  emph={rsp,rdx,rax,rbx,rbp,rsi,rdi,rcx,r8,r9,r10,r11,r12,r13,r14,r15},
  emphstyle=\color{green},
  emph={[2]cr3},
  emphstyle={[2]\color{violet}},
  morecomment=[l]{;;}
}
% Hack for Colin to edit in low light: https://tex.stackexchange.com/questions/40495/invert-background-and-text-colours-across-whole-document-with-pdflatex
%Commands definitions
\newcommand{\setbackgroundcolour}{\pagecolor[rgb]{0.19,0.19,0.19}}  
\newcommand{\settextcolour}{\color[rgb]{0.77,0.77,0.77}}    
\newcommand{\invertbackgroundtext}{\setbackgroundcolour\settextcolour}
%Command execution. 
%If this line is commented, then the appearance remains as usual.
%\invertbackgroundtext
\usepackage{wrapfig}
\newtheorem*{remark}{Remark}

% Packages.
\usepackage{amsmath}

% If we need multiple references to a footnote, via \cref:
\usepackage{cleveref}
\crefformat{footnote}{#2\footnotemark[#1]#3}
\usepackage[T1]{fontenc}
\usepackage{fontawesome}
\usepackage[utf8]{inputenc}
\usepackage{iris}
\usepackage{mathpartir}
\renewcommand{\TirNameStyle}[1]{\hypertarget{#1}{\textsc{#1}}}
\newcommand{\RULE}[1]{\hyperlink{#1}{\textsc{#1}}\xspace}
\usepackage{xspace}
\input{macros}
\newcommand{\mytodo}[1]{% <==========================================
  \todo[linecolor=white, bordercolor=white, textcolor=white]{#1}%
}

@@

\input{intro.tex}
%\input{motivation.tex}
\input{background.tex}
\input{syntaxsemantics.tex}
\input{programlogic.tex}
%\input{soundness.tex}
\input{examples.tex}
%\input{discussion.tex} % not necessary
\input{relwork.tex}
\section{Conclusions}
This paper advances the state of the art in formal verification of programs subject to address translation
on hardware using virtual memory. We proposed to treat assertions about virtual
memory locations explicitly as assertions in a modal logic, where the notion of context
is the choice of current address space based on the page table root installed on the CPU.
We improved the modularity of our virtual address translation to allow page table modifications that
preserve mappings without collecting all affected virtual points-to assertions.
To make specifications of code involving other address spaces cleaner, we borrowed the idea of
modalities which explicitly name the conditions under which they are true from hybrid logics.
We implemented these ideas in a derived separation logic within Iris, and proved soundness of
the rules for essential memory- and address-space-change-related x86-64 instructions 
sound against a hardware model of 64-bit 4-level address translation.
Finally, we used our rules to verify the correctness of key VMM instruction sequences,
including giving the first assembly-level proof of correctness in a program logic for a change
of address space, against a specification which clearly identifies which assertions hold in which address space.

% In the future we plan to 
% extend our work in multiple directions, such as generalizing our assertions and proof rules
% to support large-page functionality common to virtual memory subsystems (superpages, hugepages),
% implementing

%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{vmm}
